---
title: "simpleRegress"
author: "Misty O'Brien"
date: "Spring 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
bears$SEX <- as.factor(bears$SEX)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```
## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).

```{r}
AgeModel <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(AgeModel)
summary.lm(AgeModel)
```

The residual standard error is 7.511 and the adjusted r-squared is 0.5075. Since the residual standard error is high and the adjusted r-squared is low, this is not the best model to predict. 

```{r}
HeadLengthModel <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(HeadLengthModel)
summary.lm(HeadLengthModel)
```

The residual standard error is 4.235 and the adjusted r-squared is 0.8434 Since the residual standard error is low and the adjusted r-squared is high, this is the best model to predict so far.

```{r}
HeadWidthModel <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(HeadWidthModel)
summary.lm(HeadWidthModel)
```
The residual standard error is 7.32 and the adjusted r-squared is 0.5371. Since the residual standard error is high and the adjusted r-squared is low, this is not the best model to predict

```{r}
ChestModel <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(ChestModel)
summary.lm(ChestModel)
```

```{r}
WeightModel <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(WeightModel)
summary.lm(WeightModel)
```

# Summary Statement  

We created six different models from the bears data to predict length from other measurements. We found that the best model for this was LENGTH~HEADLEN. This model had the residual standard error of 4.235 and the adjusted r-squared of 0.8434. That shows that there is 84% of error not happening and that this model is the one with the least errors. Since the residual standard error is low and the adjusted r-squared is high, this is good and shows that its the best model. 

The next best model is the LENGTH~CHEST. This model had a residual standard error of 4.939 and an adjusted R-squared of 0.7869. This model was close second with the   residual standard error just slightly higher and the adjusted R-squared just slightly lower than the length model
